Figures from: Impacts of host availability and temperature on mosquito-borne parasite transmission

### 2023-12-19 manuscript-figures.Rmd
ensure captions are consistent with ms

```{r fig-1_contact_comparison, dev = c('pdf','svg', 'png'), out.width='100%', fig.align='center', fig.height=5, fig.width=7, fig.align='center', fig.cap='\\label{fig:contact_comparison} A comparison of the Chitnis dynamic contact rate and the standard Ross-Macdonald contact rate models. In the left columns, mosquito population is held constant while in the right columns host population density is held constant. Upper panel: In the Ross-Macdonald model, the mosquito biting rate is constant with respect to both host and mosquito population density (a and b, blue lines) while in the Chitnis model mosquito biting increases with host population density (a, red curve) and decreases with mosquito population density (b, red curve). Lower panels: The relationships between the number of bites that a host receives per day and host and mosquito densities diverge at low host population densities (c) and, to a lesser extent, at high mosquito densities (d). For plots (a) and (c), mosquito population density is fixed at 1000 individuals per km2. For plots (b) and (d), host population density is fixed at 10 individuals per square kilometer. Maximum mosquito biting rate is 0.45 bites per mosquito per day and maximum host biting rate is 100 bites per host per day. Mosquito and host population densities are plotted on a logarithmic scale. '}
# resolution setting
resolution <- 1E2

@@

```{r fig-3_MosquitoThermalTraits, dev = c('pdf','svg', 'png'),  out.width='100%', fig.height=6, fig.width=9, fig.align='center', fig.cap='\\label{fig:MosquitoThermalTraits} Thermal performance curves (TPCs) for the mosquito and parasite traits used in our model. Curves are shown in color for each of the five systems we considered. Temperatures vary from 10\textdegree C to 35\textdegree C to encompass the full thermal niche of all mosquito species considered in this study (defined as the range of temperatures at which all traits for a given species are positive). Each trait is a unimodal function of temperature except for the lifespan of \\textit{Culex quinquefasciatus} (d). Recruitment rate (a) is derived from a sub-model of immature mosquito population dynamics in equations  (2.6) and (2.7. The \\textit{Aedes aegypti} and DENV and \\textit{Ae. aegypti} and ZIKV recruitment rate (a) and maximum biting frequency (b) curves overlap because infection is assumed to not affect these traits (but see Tesla et al. (2018)). All functions and parameters for generating these curves are derived from Mordecai et al., 2013, 2017, 2019; Tesla et al., 2018; Shocket et al., 2020. Shown here are the median values of the mosquito traits with respect to their TPC hyperparameter posterior distributions. A full description of how these curves were obtained is available in Appendix S2: Section S1.'}
## Set up plot dataframe
MosqThermTraits_df <- median.Vec %>%
  # Mortality rate
  mutate(muV = 1 / lf) %>%
  # Recruitment rate
  mutate(lambdaV = V0 * muV) %>%
  # Prob. of surviving EIP
  mutate(P_EIP = exp(-muV / etaV)) %>%
  # Prob. of larval survival
  # Select relevant traits and parameters
  dplyr::select(Temperature, system_ID, sigmaV, sigmaV_f, rhoL, lf, deltaL, P_EIP, betaV) %>%
  # Turn all columns except Temperature and system_ID into variables
  melt(id = c("Temperature", "system_ID")) %>%
  mutate(variable = factor(variable,
                           levels = c("lf", "sigmaV", "sigmaV_f", "rhoL", "deltaL", "P_EIP", "betaV"),
                           labels = c(
                             TeX("(a) Adult lifespan: $1/\\mu_V$ (days)"),
                             TeX("(b) Biting frequency: $\\sigma_V$ (day$^{-1}$)"),
                             TeX("(c) Eggs per female per day: $\\sigma_V f$"),
                             TeX("(d) Mosquito development rate: $\\rho_L$ (day$^{-1}$)"),
                             TeX("(e) Prob. of immature mosquito survival: $\\delta_L$"),
                             TeX("(f) Probability of surviving EIP: $\\theta_V$"),
                             TeX("(g) Vector competence: $\\beta_V$")
                           )
  )) %>%
  group_by(Temperature, system_ID)

@@

```{r fig-4_R0_TPCs, dev = c('pdf','svg', 'png'),  message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:R0_plots} $\\mathcal{R}_0$ as a function of temperature has the usual properties of a thermal performance curve (TPC): it is positive on a single bounded interval and has a unique maximum value. The shape of the $\\mathcal{R}_0$ TPC depends on vertebrate host availability and the mosquito-parasite system. The shape of the thermal response of $\\mathcal{R}_0$ derived from the Ross-Macdonald model (dotted black curves) is independent of vertebrate host population density while those derived from the Chitnis dynamic model are dependent on vertebrate host population density (solid lines in color). Shown here are the median values of $\\mathcal{R}_0$ with respect to the mosquito trait TPC posterior distributions. $\\mathcal{R}_0$ is normalized by dividing by its maximum value in each system, for the given value of biting tolerance and vertebrate host population density. Vertebrate host population density is varied on a log$_{10}$ scale.'}
# Plot R0 vs. temperature curves. One plot for each system
# One curve for Ross-Macdonald, a few for Chitnis

@@

```{r fig-5_Topt, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt} The thermal optimum for transmission, $T_{\\mathrm{opt}}$, may be sensitive to vertebrate host population density and biting tolerance. Here, $T_{\\mathrm{opt}}$ is plotted as a function of vertebrate host population density and biting tolerance for each of the five systems. In the Ross-Macdonald model, where biting tolerance is assumed to be infinite, $T_{\\mathrm{opt}}$ is independent of vertebrate host population density (upper bars in plots). But when biting tolerance is limited, $T_{\\mathrm{opt}}$ can vary as much as 5\textdegree C (d). $T_{\\mathrm{opt}}$ may increase (d) or decrease (a, b, c, e) as vertebrate host population density or biting tolerance is increased. Shown here are the median values of $T_{\\mathrm{opt}}$ with respect to the mosquito trait TPC posterior distributions. Both biting tolerance and vertebrate host population density are varied on a $\\log_{10}$ scale. $\\mathcal{R}_0$ is greater than one in the region contained within the black $\\mathcal{R}_0=1$ curve. All remaining parameters are given in Table 1 and Figure 3. '}
# Shows how the transmission thermal optimum depends on vertebrate host availability (density and biting tolerance) in each of the systems we consider.

@@

```{r fig-6_CT_Width, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_Width} The parasite population thermal niche, the range of temperatures at which parasite transmission is expected to persist, is widest when biting tolerance is high and vertebrate host population density is low. As vertebrate host population density approaches an upper threshold, the width of the parasite population thermal niche shrinks to zero. At low levels of biting tolerance, there is also a lower threshold to vertebrate host population density. Shown here are the median values of $\\textrm{CT}_{\\textrm{width}}$ with respect to the mosquito trait TPC posterior distributions. Vertebrate host population density and biting tolerance are both plotted on $\\log_{10}$ scales. Biting tolerance is varied from one to one thousand bites per day, with an additional point at infinity to illustrate the case where biting tolerance is unlimited (the Ross-Macdonald model). Both biting tolerance and vertebrate host population density are varied on a $\\log_{10}$ scale. All remaining parameters are given in Table 1 and Figure 3.'}
# Shows how the parasite thermal niche (temperatures between CT_min and CT_max)
# depends on vertebrate host availability (density and biting tolerance) in each
# of the focal systems.

@@

```{r fig-7_explain_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:dR0dvar_comp} Comparison of the local sensitivities of $\\mathcal{R}_0$ to mosquito traits across temperatures for two values of $\\sigma_H$. Generally, $\\mathcal{R}_0$ is most sensitive to biting frequency, the probability of surviving the extrinsic incubation period, and vector competence. Shown here are the median values of the derivative of $\\mathcal{R}_0$ with respect to each focal mosquito trait across the mosquito trait TPC hyperparameter posterior distributions. Sensitivity values are shown on a logarithmic scale with solid lines indicating positive values and dotted lines indicating negative. Temperatures are restricted to those at which the median of $\\mathcal{R}_0$ is positive for the given system.'}

@@

```{r fig-S0_MosquitoDensity, dev = c('pdf','svg', 'png'),  out.width='100%', fig.height=6, fig.width=9, fig.align='center', fig.cap='\\label{fig:MosquitoDensity} Thermal performance curves (TPCs) for the mosquito and parasite traits used in our model. Curves are shown in color for each of the five systems we considered. Temperatures vary from 10\textdegree C to 35\textdegree C to encompass the full thermal niche of all mosquito species considered in this study (defined as the range of temperatures at which all traits for a given species are positive). Each trait is a unimodal function of temperature except for the lifespan of \\textit{Culex quinquefasciatus} (d). Recruitment rate (a) is derived from a sub-model of immature mosquito population dynamics in equations  (2.6) and (2.7. The \\textit{Aedes aegypti} and DENV and \\textit{Ae. aegypti} and ZIKV recruitment rate (a) and maximum biting frequency (b) curves overlap because infection is assumed to not affect these traits (but see Tesla et al. (2018)). All functions and parameters for generating these curves are derived from Mordecai et al., 2013, 2017, 2019; Tesla et al., 2018; Shocket et al., 2020. Shown here are the median values of the mosquito traits with respect to their TPC hyperparameter posterior distributions. A full description of how these curves were obtained is available in Appendix S2: Section S1.'}

@@

```{r fig-S1_R0_heatmap_plots_all, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=11, fig.align='center', fig.cap='\\label{fig:R0_heatmap_plots_all} The transmission thermal optimum, $T_{\\textrm{opt}}$, is a sigmoidal function of vertebrate host population density. The variation in $T_{\\textrm{opt}}$ varies substantially between the five systems, with \\textit{Anopheles} spp. and \\textit{P. falciparum} (d) having an over 5\textdegree C change in $T_{\\textrm{opt}}$ as vertebrate host population density is increased. Note that the thermal optimum can be an increasing (d) or decreasing (a, b, c, e) function of vertebrate host population density. Shown here are the median values of $T_{\\textrm{opt}}$ with respect to mosquito trait TPC hyperparameter posterior distributions. Vertebrate host population density is varied on a $\\log_{10}$ scale. All remaining parameters are given in Table 1 and Figure 3.'}

@@

```{r fig-S2_Topt_plots_varyKH, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_plots_varyKH} The transmission thermal optimum, $T_{\\textrm{opt}}$, is a sigmoidal function of vertebrate host population density. The variation in $T_{\\textrm{opt}}$ varies substantially between the five systems, with \\textit{Anopheles} spp. and \\textit{P. falciparum} (d) having an over 5\textdegree C change in $T_{\\textrm{opt}}$ as vertebrate host population density is increased. Note that the thermal optimum can be an increasing (d) or decreasing (a, b, c, e) function of vertebrate host population density. Shown here are the median values of $T_{\\textrm{opt}}$ with respect to mosquito trait TPC hyperparameter posterior distributions. Vertebrate host population density is varied on a $\\log_{10}$ scale. All remaining parameters are given in Table 1 and Figure 3.'}

@@

```{r fig-S4_CTmin, dev = c('pdf','svg', 'png'), dev="cairo_pdf", message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CTmin} The critical thermal minimum for parasite population persistence as a function of vertebrate host population density and biting tolerance. Grey areas indicate combinations of vertebrate host population density and biting tolerance in which the parasite population is expected to not persist at any temperature. In the region where vertebrate host population density and biting tolerance are both less than 10, critical thermal minima and maxima are very similar (see Figure 6 in the main text). Shown here are the median values of the critical thermal minimum with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited. '}
#   x    = biting tolerance (with a point at infinity)
#   y    = vertebrate host population density
# colour = parasite critical thermal minimum (CTmin)
# Facets = systems

@@

```{r fig-S5_CTmax, dev = c('pdf','svg', 'png'), dev="cairo_pdf", message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CTmax} The critical thermal maximum for parasite population persistence as a function of vertebrate host population density and biting tolerance. Grey areas indicate combinations of vertebrate host population density and biting tolerance in which the parasite population is expected to not persist at any temperature. In the region where vertebrate host population density and biting tolerance are both less than 10, critical thermal minima and maxima are very similar (see Figure 6 in the main text). Shown here are the median values of the critical thermal maximum with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited. '}
#   x    = biting tolerance (with a point at infinity)
#   y    = vertebrate host population density
# colour = parasite critical thermal maximum (CTmax)
# Facets = systems

@@

```{r fig-S6_CT_Range, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_Range} The critical thermal minimum ($\\textrm{CT}_{\\min}$, blue curves) and maximum ($\\textrm{CT}_{\\max}$, red curves) for parasite population persistence depend on vertebrate host availability, even when vertebrate hosts tolerate any level of mosquito biting (dotted curves). Transmission is sustained ($\\mathcal{R}_0$>1) in the regions enclosed by the red and blue curves. $\\textrm{CT}_{\\min}$ is a convex function of vertebrate host population density while $\\textrm{CT}_{\\max}$ is concave. The points where the red and blue curves touch indicate thresholds of vertebrate host population density beyond which parasite population persistence is no longer possible at any temperature (i.e. $\\mathcal{R}_0$(T)<1) (see Figure 6 in the main text). Shown here are the median values of the critical thermal minimum, maximum, and mosquito thermal tolerance niche with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited. All remaining parameters are given in Table 1 and Figure 2.'}
# x-axis = vertebrate host population density
# y-axis = Temperature
# linetype = biting tolerance
# colour = CT_min (blue), CT_max (red)

@@

```{r fig-S7_CT_distributions, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_dists} The distributions of the critical thermal minimum and maximum and thermal optimum vary across transmission systems and levels of vertebrate host availability. Distributions for the limited biting tolerance case (solid lines) and unlimited biting tolerance case (dashed lines) are shown above and below the vertebrate host population density levels, respectively. As vertebrate host population density is increased, the distributions become approximately equal, with the critical thermal minimum and maximum converging to each other at sufficiently high population densities. The \\textit{An. gambiae} and \\textit{P. falciparum} system shows the most drastic differences between the limited and unlimited biting tolerance cases at low vertebrate host population density levels. Vertebrate host population density is varied on a $\\log_{10}$ scale. '}

@@

```{r fig-S8_R0_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:dR0dvar_unc} The local sensitivity of $\\mathcal{R}_0$ to mosquito traits across temperatures for two values of $\\sigma_H$ and $K_H$. Generally, $\\mathcal{R}_0$ is most sensitive to biting frequency, the probability of surviving the extrinsic incubation period, and vector competence. Shown here are the median values of the derivative of $\\mathcal{R}_0$ with respect to each focal mosquito trait across the mosquito trait TPC hyperparameter posterior distributions. Sensitivity values are shown on a logarithmic scale with solid lines indicating positive values and dotted lines indicating negative. Temperatures are restricted to those at which the median of $\\mathcal{R}_0$ is positive for the given system. '}

@@

```{r fig-S10_Topt_uncertainty, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_unc} The uncertainty of $T_{\\textrm{opt}}$ to mosquito traits across vertebrate host population densities as measured by the change in the 95\\% highest posterior density interval width when $\\sigma_H$ is 100 bites per vertebrate host per day. At high levels of vertebrate host population density, the effect of vertebrate host traits diminishes and the relative sensitivities of $T_{\\textrm{opt}}$ to mosquito traits stabilize. Generally, adult lifespan is an influential parameter and, to lesser extents, vector competence and the probability of surviving the extrinsic incubation period (EIP). Eggs per female per day tends to be most important when vertebrate host population density is low, though this not the case for (e) \\textit{Cx. quinquefasciatus} and WNV.Immature development rate and the probability of immature survival are never very influential parameters. Vertebrate host population density is varied on a $\\log_{10}$ scale. '}

>>>>

```{r fig-1_contact_comparison, dev = c('pdf','svg', 'png'), out.width='100%', fig.align='center', fig.height=5, fig.width=7, fig.align='center', fig.cap='\\label{fig:contact_comparison} A comparison of the Chitnis dynamic contact rate and the standard Ross-Macdonald contact rate models. In the left columns, mosquito population is held constant while in the right columns host population density is held constant. Upper panel: In the Ross-Macdonald model, the mosquito biting rate is constant with respect to both host and mosquito population density (a and b, blue lines) while in the Chitnis model mosquito biting increases with host population density (a, red curve) and decreases with mosquito population density (b, red curve). Lower panels: The relationships between the number of bites that a host receives per day and host and mosquito densities diverge at low host population densities (c) and, to a lesser extent, at high mosquito densities (d). For plots (a) and (c), mosquito population density is fixed at 1000 individuals per square kilometer. For plots (b) and (d), host population density is fixed at 10 individuals per square kilometer. Maximum mosquito biting rate is 0.45 bites per mosquito per day and maximum host biting rate is 100 bites per host per day. Mosquito and host population densities are plotted on a logarithmic scale.'}

# resolution setting
resolution <- 1E2

@@

```{r fig-3_MosquitoThermalTraits, dev = c('pdf','svg', 'png'),  out.width='100%', fig.height=6, fig.width=9, fig.align='center', fig.cap='\\label{fig:MosquitoThermalTraits} Thermal performance curves (TPCs) for the mosquito and parasite traits used in our model. Curves are shown in color for each of the five systems we considered. Temperatures vary from 10\\textdegree C to 35\\textdegree C to encompass the full thermal niche of all mosquito species considered in this study (defined as the range of temperatures at which all traits for a given species are positive). Each trait is a unimodal function of temperature except for the lifespan of \\textit{Culex quinquefasciatus} (d). Recruitment rate (a) is derived from a sub-model of immature mosquito population dynamics described in Appendix S2. The \textit{Aedes aegypti} and DENV and \textit{Ae. aegypti} and ZIKV recruitment rate (a) and maximum biting frequency (b) curves overlap because infection is assumed to not affect these traits (but see Tesla et al. (2018)). All functions and parameters for generating these curves are derived from Mordecai et al. (2013, 2017, 2019); Tesla et al. (2018); Shocket et al. (2020); Villena et al. (2022). Shown here are the median values of the mosquito traits with respect to their TPC hyperparameter posterior distributions. A full description of how these curves were obtained is available in Appendix S2: Section S1.'}

## Set up plot dataframe
MosqThermTraits_df <- median.Vec %>%
  # Mortality rate
  mutate(muV = 1 / lf) %>%
  # Recruitment rate
  mutate(lambdaV = V0 * muV) %>%
  # Prob. of surviving EIP
  mutate(P_EIP = exp(-muV / etaV)) %>%
  # Prob. of larval survival
  # Select relevant traits and parameters
  dplyr::select(Temperature, system_ID, sigmaV, sigmaV_f, rhoL, lf, deltaL, P_EIP, betaV) %>%
  # Turn all columns except Temperature and system_ID into variables
  melt(id = c("Temperature", "system_ID")) %>%
  mutate(variable = factor(variable,
                           levels = c("lf", "sigmaV", "sigmaV_f", "rhoL", "deltaL", "P_EIP", "betaV"),
                           labels = c(
                             TeX("(a) Adult lifespan: $1/\\mu_V$ (days)"),
                             TeX("(b) Biting frequency: $\\sigma_V$ (day$^{-1}$)"),
                             TeX("(c) Eggs per female per day: $\\sigma_V f$"),
                             TeX("(d) Mosquito development rate: $\\rho_L$ (day$^{-1}$)"),
                             TeX("(e) Prob. of immature mosquito survival: $\\delta_L$"),
                             TeX("(f) Probability of surviving EIP: $\\theta_V$"),
                             TeX("(g) Vector competence: $\\beta_V$")
                           )
  )) %>%
  group_by(Temperature, system_ID)

@@

```{r fig-4_R0_TPCs, dev = c('pdf','svg', 'png'),  message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:R0_plots} $\\mathcal{R}_0$ as a function of temperature has the usual properties of a thermal performance curve (TPC): it is positive on a single bounded interval and has a unique maximum value. The shape of the $\\mathcal{R}_0$ TPCdepends on vertebrate host availability and the mosquito-parasite system. The shape of the thermal response of $\\mathcal{R}_0$ derived from the Ross-Macdonald model (dotted black curves) is independent of vertebrate host population density while those derived from the Chitnis dynamic model are dependent on vertebrate host population density (solid lines in color). Shown here are the median values of $\\mathcal{R}_0$ with respect to the mosquito trait TPC posterior distributions. $\\mathcal{R}_0$ is normalized by dividing by its maximum value in each system, for the given value of biting tolerance and vertebrate host population density. Vertebrate host population density is varied on a log$_{10}$ scale.'}

# Plot R0 vs. temperature curves. One plot for each system
# One curve for Ross-Macdonald, a few for Chitnis

@@

```{r fig-5_Topt, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt} The thermal optimum for transmission, $T_{\\mathrm{opt}}$, may be sensitive to vertebrate host population density and biting tolerance. Here, $T_{\\mathrm{opt}}$ is plotted as a function of vertebrate host population density and biting tolerance for each of the five systems. In the Ross-Macdonald model, where biting tolerance is assumed to be infinite, $T_{\\mathrm{opt}}$ is independent of vertebrate host population density (upper bars in plots). But when biting tolerance is limited, $T_{\\mathrm{opt}}$ can vary as much as 5\\textdegree C (d). $T_{\\mathrm{opt}}$ may increase (d) or decrease (a, b, c, e) as vertebrate host population density or biting tolerance is increased. Shown here are the median values of $T_{\\mathrm{opt}}$ with respect to the mosquito trait TPC posterior distributions. Both biting tolerance and vertebrate host population density are varied on a $\\log_{10}$ scale. $\\mathcal{R}_0$ is greater than one in the region contained within the black $\\mathcal{R}_0=1$ curve. All remaining parameters are given in Table 1 and Figure 3.'}

# Shows how the transmission thermal optimum depends on vertebrate host availability (density and biting tolerance) in each of the systems we consider.

@@

```{r fig-6_CT_Width, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_Width} The parasite population thermal niche, the range of temperatures at which parasite transmission is expected to persist, is widest when biting tolerance is high and vertebrate host population density is low. As vertebrate host population density approaches an upper threshold, the width of the parasite population thermal niche shrinks to zero. At low levels of biting tolerance, there is also a lower threshold to vertebrate host population density. Shown here are the median values of $\\textrm{CT}_{\\textrm{width}}$ with respect to the mosquito trait TPC posterior distributions. Vertebrate host population density and biting tolerance are both plotted on $\\log_{10}$ scales. Biting tolerance is varied from one to one thousand bites per day, with an additional point at infinity to illustrate the case where biting tolerance is unlimited (the Ross-Macdonald model). Both biting tolerance and vertebrate host population density are varied on a $\\log_{10}$ scale. All remaining parameters are given in Table 1 and Figure 3.'}

# Shows how the parasite thermal niche (temperatures between CT_min and CT_max)
# depends on vertebrate host availability (density and biting tolerance) in each
# of the focal systems.

@@

```{r fig-7_explain_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:dR0dvar_comp} The change in the contributions of the temperature-dependent mosquito traits to the overall temperature-sensitivity of $\\mathcal{R}_0$ as vertebrate host population density is increased from $1$ to $100$ individuals per hectare. There is a fairly consistent pattern in the change in importance for four out of the five systems.But, unlike the others, the \\textit{Anopheles gambiae} and \\textit{Plasmodium falciparum} system displays large changes in the contributions of biting frequency and parasite development rate at low temperatures and relatively small changes for adult mortality rate at warmer temperatures. This indicates that, through the incorporation of limited biting tolerance, a change in vertebrate host availability shifts the importance of certain temperature-dependent mosquito traits in determining the temperature-dependence of $\\mathcal{R}_0$. Shown here are the median values of $\\bar{\\mathcal{S}_p}(T, 100, 10) - \\bar{\\mathcal{S}_p}(T, 1, 10)$ as described by equation (2.4).'}

@@

```{r fig-S0_MosquitoDensity, dev = c('pdf','svg', 'png'),  out.width='100%', fig.height=6, fig.width=9, fig.align='center', fig.cap='\\label{fig:MosquitoDensity} Equilibrium mosquito population density, $V^*(T)$, as a function of temperature. Shown here are the median values of $V^*(T)$ with respect to the mosquito trait TPC hyperparameter posterior distributions. Because infection was not assumed to affect life history traits, the \\textit{Aedes aegypti} and DENV and \\textit{Ae. aegypti} and ZIKV curves overlap. As shown in Equation (2.1), $V^*(T)$ is a function of the immature mosquito development rate, probability of immature mosquito survival to adulthood, eggs per female per day, the adult mortality rate, and the carrying capacity for larval mosquitoes. This carrying capacity was assumed to be temperature-independent and set to 300, which resulted in a plausible range of mosquito population densities for the five systems. '}

@@

```{r fig-S1_R0_heatmap_plots_all, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=11, fig.align='center', fig.cap='\\label{fig:R0_heatmap_plots_all} $\\mathcal{R}_0$ as a function of temperature and biting tolerance  ($\\sigma_H$) across three levels of vertebrate host population density: $0.1$, $1$, and $10$ vertebrate hosts per hectare. Shown here are the median values of $\\mathcal{R}_0$ with respect to mosquito trait TPC hyperparameter posterior distributions. Vertebrate host biting tolerance is varied on a $\\log_{10}$ scale with a break to indicate values of $\\mathcal{R}_0$ when biting tolerance is unlimited.'}

@@

```{r fig-S2_Topt_plots_varyKH, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_plots_varyKH} The transmission thermal optimum, $T_{\\textrm{opt}}$, is a sigmoidal function of vertebrate host population density. The variation in $T_{\\textrm{opt}}$ varies substantially between the five systems, with \\textit{Anopheles gambiae} and \\textit{Plasmodium falciparum} (d) having an over 5\\textdegree C change in $T_{\\textrm{opt}}$ as vertebrate host population density is increased. Note that the thermal optimum can be an increasing (d) or decreasing (a, b, c, e) function of vertebrate host population density. Shown here are the median values of $T_{\\textrm{opt}}$ with respect to mosquito trait TPC hyperparameter posterior distributions. Vertebrate host population density is varied on a $\\log_{10}$ scale. All remaining parameters are given in Table 1 and Figure 3.'}

@@

```{r fig-S4_CTmin, dev = c('pdf','svg', 'png'), dev="cairo_pdf", message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CTmin} The critical thermal minimum for parasite population persistence as a function of vertebrate host population density and biting tolerance. Grey areas indicate combinations of vertebrate host population density and biting tolerance in which the parasite population is expected to not persist at any temperature. In the region where vertebrate host population density and biting tolerance are both less than 10, critical thermal minima and maxima are very similar (see Figure 6 in the main text). Shown here are the median values of the critical thermal minimum with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited. '}

#   x    = biting tolerance (with a point at infinity)
#   y    = vertebrate host population density
# colour = parasite critical thermal minimum (CTmin)
# Facets = systems

@@

```{r fig-S5_CTmax, dev = c('pdf','svg', 'png'), dev="cairo_pdf", message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CTmax} The critical thermal maximum for parasite population persistence as a function of vertebrate host population density and biting tolerance. Grey areas indicate combinations of vertebrate host population density and biting tolerance in which the parasite population is expected to not persist at any temperature. In the region where vertebrate host population density and biting tolerance are both less than 10, critical thermal minima and maxima are very similar (see Figure 6 in the main text). Shown here are the median values of the critical thermal maximum with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited.'}

#   x    = biting tolerance (with a point at infinity)
#   y    = vertebrate host population density
# colour = parasite critical thermal maximum (CTmax)
# Facets = systems

@@

```{r fig-S6_CT_Range, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_Range} The critical thermal minimum ($\\textrm{CT}_{\\min}$, blue curves) and maximum ($\\textrm{CT}_{\\max}$, red curves) for parasite population persistence depend on vertebrate host availability, even when vertebrate hosts tolerate any level of mosquito biting (dotted curves). Transmission is sustained ($\\mathcal{R}_0$>1) in the regions enclosed by the red and blue curves. $\\textrm{CT}_{\\min}$ is a convex function of vertebrate host population density while $\\textrm{CT}_{\\max}$ is concave. The points where the red and blue curves touch indicate thresholds of vertebrate host population density beyond which parasite population persistence is no longer possible at any temperature (i.e. $\\mathcal{R}_0$(T)<1) (see Figure 6 in the main text). Shown here are the median values of the critical thermal minimum, maximum, and mosquito thermal tolerance niche with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited. All remaining parameters are given in Table 1 and Figure 2.'}

# x-axis = vertebrate host population density
# y-axis = Temperature
# linetype = biting tolerance
# colour = CT_min (blue), CT_max (red)

@@

```{r fig-S7_CT_distributions, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_dists} The distributions of the critical thermal minimum and maximum and thermal optimum vary across transmission systems and levels of vertebrate host availability. Distributions for the limited biting tolerance case (solid lines) and unlimited biting tolerance case (dashed lines) are shown above and below the vertebrate host population density levels, respectively. As vertebrate host population density is increased, the distributions become approximately equal, with the critical thermal minimum and maximum converging to each other at sufficiently high population densities. The \\textit{An. gambiae} and \\textit{P. falciparum} system shows the most drastic differences between the limited and unlimited biting tolerance cases at low vertebrate host population density levels. Vertebrate host population density is varied on a $\\log_{10}$ scale.'}

@@

```{r fig-S8_R0_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:dR0dvar_unc} The local sensitivity of $\\mathcal{R}_0$ to mosquito traits across temperatures for two values of biting tolerance ($\\sigma_H$) and vertebrate host population density ($K_H$). Overall, the sensitivity of $\\mathcal{R}_0$ to temperature decreases at higher temperatures. At high temperatures, adult mortality rate is the primary contributor to the temperature sensitivity of $\\mathcal{R}_0$ across all systems and all values of $\\sigma_H$ and $K_H$. At lower temperatures, the most important trait varies between the adult mortality rate, biting frequency, parasite development rate, or vector competence, depending on the system in question and values of the vertebrate host traits. The values shown here are the median values taken across the mosquito trait TPC hyperparameter posterior distributions of the contributions of each trait to the local sensitivity of derivative of $\\mathcal{R}_0$ to temperature: $\\frac{1}{\\mathcal{R}_0}\\frac{\\partial\\mathcal{R}_0}{\\partial p}\\frac{dp}{dT}$ where $p$ is a focal trait and $T$ is temperature. Values were scaled to span from -1 to 1 to enable comparisons across systems. Temperatures are restricted to those at which $\\mathcal{R}_0$ is positive for the given system. '}

@@

```{r fig-S10_Topt_uncertainty, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_unc} The uncertainty of $T_{\\textrm{opt}}$ to mosquito traits across vertebrate host population densities as measured by the change in the 95\\% highest posterior density interval width when $\\sigma_H$ is 100 bites per vertebrate host per day. At high levels of vertebrate host population density, the effect of vertebrate host traits diminishes and the relative sensitivities of $T_{\\textrm{opt}}$ to mosquito traits stabilize. Generally, adult lifespan is an influential parameter and, to lesser extents, vector competence and the probability of surviving the extrinsic incubation period (EIP). Eggs per female per day tends to be most important when vertebrate host population density is low, though this not the case for (e) \\textit{Cx. quinquefasciatus} and WNV. Immature development rate and the probability of immature survival are never very influential parameters. Vertebrate host population density is varied on a $\\log_{10}$ scale. '}
